Urban Heat Risk Mapping Using Multiple Point Patterns in Houston, Texas

### 2016-08-11 jrssc-mpp.tex
Provided justification for using grid cells as opposed to census tracts

\subsection{Likelihood Layer}\label{likelayer}
Let $\mathbf{s}_{pi} \in \mathcal{D}$ be the latitude-longitude coordinate
of the location for the $i^{\text{th}}$ 
individual who experienced health outcome $p$ where $i=1,2,\dots,n_p$, $p=1,2,\dots,P$, 
such that $P$ is the total number of health outcomes under consideration (in this article, $P=2$),
and $n_p$ is the total number of individuals who experienced health outcome $p$.
Likewise, let $a_{pi} \in \left\{0,1,2,\dots,100\right\}$ represent the age of the $i^{\text{th}}$ individual who
experienced health outcome $p$; 
$g_{pi} \in \left\{\text{female}, \text{male}\right\}$ represent the gender of individual $i$ who experienced health 
outcome $p$; and $e_{pi} \in \left\{\text{Asian}, \text{Black}, \text{Hispanic}, 
\text{White}\right\}$ be the race/ethnicity of individual $i$ who experienced health
outcome $p$. The upper bound of 100 for age was chosen arbitrarily but, in our data, 
there were no individuals with an age greater than 100. For other applications, this upper 
bound may need to be adjusted to account for the possibility of older ages. 
Finally, define $\mathbf{z}_{pi} = (\mathbf{s}_{pi}',a_{pi},g_{pi},e_{pi})' \in 
\mathcal{Z}$ to be the vector of realized spatial locations and marks.

@@

The population data used to calculate $E_{pk}$ was taken from the 2010 census. 
Over the area covered
by our $1428$ grid cells there are $1536$ irregularly shaped census block groups. We handle
the misalignment between the block groups and our grid cells using a method outlined
by \cite{Brunsdon2015}. In this process we calculate the intersection
between grid cells and census block groups and use this to calculate the proportion of
a grid cell that lies within a block group. If a grid cell spans multiple
census block groups, the intersection will be split between those groups; if the grid cell 
is entirely contained within a block group, then the intersection will just be the grid cell 
itself. Let $\mathcal{C}_i$ represent census block group $i$ and denote the proportion
$\omega_{ki}$ as
$$\omega_{ki} = \frac{\left| \mathcal{G}_k \cap \mathcal{C}_i \right|}
  {\left|\mathcal{C}_i\right|}.$$ 
In this equation $\left| \mathcal{G}_k \cap \mathcal{C}_i \right|$ represents
the area of the intersection between grid cell $k$ and block group $i$ and 
$\left| \mathcal{C}_i \right|$ is the area of block group $i$. 
Defining $\text{CensPop}_i$ as the population count in $\mathcal{C}_i$ allows us to approximate the population count
within grid cell $k$ as $$\text{Pop}_k = \sum_{i=1}^{N} \omega_{ki}\cdot \text{CensPop}_i,$$
where $N$ is the total
number of census block groups that intersect with grid cell $k$. Therefore, population within
a given block group is distributed among intersecting grid cells based on the proportion
of the group contained within a given grid cell. One obvious weakness of this 
method is that it assumes that the population within
a census block group is distributed equally throughout that group, which is almost
certainly not the case; however, due to restrictions on census data, demographic information (which we use elsewhere in the model) is not available at a finer resolution. As a result, we feel this approach is justified in order to keep all census data incorporated in the model on the same scale.

@@

\subsection{Hyperprior Layer}
At the hyperior prior layer of the model, we outline our prior assumptions for the model 
parameters in Sections \ref{likelayer} and \ref{process}. Because $\boldsymbol{\alpha}_p = 
(\alpha_p(0),\dots,\alpha_p(100))'$ are probabilities where $\sum_a \alpha_p(a) = 1$, 
we assume $\boldsymbol{\alpha}_{p} \sim \text{Dir}(2\times\mathbf{1}_{100})$ for all $p$ where 
$\text{Dir}(\cdot)$ denotes the Dirichlet distribution and $\mathbf{1}_{100}$ is a $1\times 100$ vector of ones. This prior specification is both non-informative and conjugate. 
Using the likelihood in \eqref{likelihood}, note that the posterior distribution for 
$\boldsymbol{\alpha}_p$ is
\begin{align}
\text{post}(\boldsymbol{\alpha}_p) &\propto \left[\prod_{a=0}^{100} \alpha_p^{n_{pa}}(a)\right]
\left[\prod_{a=0}^{100}\alpha_p^{2-1}(a)\right] \notag \\
&= \prod_{a=0}^{100} \alpha_p^{n_{pa}+2-1}(a),
\label{apost}
\end{align} 
where $n_{pa}$ is the number of individuals who experienced health outcome $p$ and were age 
$a$. The form of the posterior distribution in \eqref{apost} is $\text{Dir}(n_{p0}+2,n_{p1}+2,
\dots,n_{p100}+2)$. Similar prior distributions were assumed for $\{\gamma_p(g)\}_{p,g}$ 
and $\{\xi_p(e)\}_{p,e}$.

>>>>

\subsection{Likelihood Layer}\label{likelayer}
Let $\mathbf{s}_{pi} \in \mathcal{D}$ be the latitude-longitude coordinate
of the location for the $i^{\text{th}}$ 
individual who experienced health outcome $p$ where $i=1,2,\dots,n_p$, $p=1,2,\dots,P$, 
such that $P$ is the total number of health outcomes under consideration (in this article, $P=2$),
and $n_p$ is the total number of individuals who experienced health outcome $p$.
Likewise, let $a_{pi} \in \left\{0,1,2,\dots,113\right\}$ represent the age of the $i^{\text{th}}$ individual who
experienced health outcome $p$; 
$g_{pi} \in \left\{\text{female}, \text{male}\right\}$ represent the gender of individual $i$ who experienced health 
outcome $p$; and $e_{pi} \in \left\{\text{Asian}, \text{Black}, \text{Hispanic}, 
\text{White}\right\}$ be the race/ethnicity of individual $i$ who experienced health
outcome $p$. The upper bound of 113 for age was chosen arbitrarily but, in our data, 
there were no individuals with an age greater than 113. For other applications, this upper 
bound may need to be adjusted to account for the possibility of older ages. 
Finally, define $\mathbf{z}_{pi} = (\mathbf{s}_{pi}',a_{pi},g_{pi},e_{pi})' \in 
\mathcal{Z}$ to be the vector of realized spatial locations and marks.

@@

The population data used to calculate $E_{pk}$ was taken from the 2000 census. 
Over the area covered
by our $1428$ grid cells there are $1461$ irregularly shaped census block groups. We handle
the misalignment between the block groups and our grid cells using a method outlined
by \cite{Brunsdon2015}. In this process we calculate the intersection
between grid cells and census block groups and use this to calculate the proportion of
a grid cell that lies within a block group. If a grid cell spans multiple
census block groups, the intersection will be split between those groups; if the grid cell 
is entirely contained within a block group, then the intersection will just be the grid cell 
itself. Let $\mathcal{C}_i$ represent census block group $i$ and denote the proportion
$\omega_{ki}$ as
$$\omega_{ki} = \frac{\left| \mathcal{G}_k \cap \mathcal{C}_i \right|}
  {\left|\mathcal{C}_i\right|}.$$ 
In this equation $\left| \mathcal{G}_k \cap \mathcal{C}_i \right|$ represents
the area of the intersection between grid cell $k$ and block group $i$ and 
$\left| \mathcal{C}_i \right|$ is the area of block group $i$. 
Defining $\text{CensPop}_i$ as the population count in $\mathcal{C}_i$ allows us to approximate the population count
within grid cell $k$ as $$\text{Pop}_k = \sum_{i=1}^{N} \omega_{ki}\cdot \text{CensPop}_i,$$
where $N$ is the total
number of census block groups that intersect with grid cell $k$. Therefore, population within
a given block group is distributed among intersecting grid cells based on the proportion
of the group contained within a given grid cell. One weakness of this 
method is that it assumes that the population within
a census block group is distributed equally throughout that group, which is almost
certainly not the case. Unfortunately, due to restrictions on census data, demographic information (which we use elsewhere in the model) is not available at a finer resolution. As a result, we feel this approach is justified in order to keep all census data incorporated in the model on the same scale.

A potential solution to avoid realigning the census data to the grid cells is to let the census block groups equal the $\mathcal{G}_k$ denoted in \eqref{lambda} rather than the grid cells. However, the heat information used in this research (described in detail in Section \ref{sec:temp_data}) is provided at a 1-km$^2$ resolution and these grid cells were chosen to align exactly with the available heat data. Using the block groups in place of the grid cells would necessitate aligning the heat information to the census block group areas. Because heat is our primary variable of interest we prefer realigning the demographic information to the grid associated with the heat data as opposed to the alternative. Therefore, we proceed having aligned the population data as described.

@@

\subsection{Hyperprior Layer}
At the hyperior prior layer of the model, we outline our prior assumptions for the model 
parameters in Sections \ref{likelayer} and \ref{process}. Because $\boldsymbol{\alpha}_p = 
(\alpha_p(0),\dots,\alpha_p(100))'$ are probabilities where $\sum_a \alpha_p(a) = 1$, 
we assume $\boldsymbol{\alpha}_{p} \sim \text{Dir}(\mathbf{1}_{100})$ for all $p$ where 
$\text{Dir}(\cdot)$ denotes the Dirichlet distribution and $\mathbf{1}_{100}$ is a $1\times 100$ vector of ones. This prior specification is both non-informative and conjugate. 
Using the likelihood in \eqref{likelihood}, note that the posterior distribution for 
$\boldsymbol{\alpha}_p$ is
\begin{align}
\text{post}(\boldsymbol{\alpha}_p) &\propto \left[\prod_{a=0}^{100} \alpha_p^{n_{pa}}(a)\right]
\left[\prod_{a=0}^{100}\alpha_p^{2-1}(a)\right] \notag \\
&= \prod_{a=0}^{100} \alpha_p^{n_{pa}+2-1}(a),
\label{apost}
\end{align} 
where $n_{pa}$ is the number of individuals who experienced health outcome $p$ and were age 
$a$. The form of the posterior distribution in \eqref{apost} is $\text{Dir}(n_{p0}+2,n_{p1}+2,
\dots,n_{p100}+2)$. Similar prior distributions were assumed for $\{\gamma_p(g)\}_{p,g}$ 
and $\{\xi_p(e)\}_{p,e}$.
